---
title: 'Ridehail Notebooks: Sequences'
output:
  html_document: 
    toc: yes
    toc_depth: 2
---

# Toronto, Feb 6, 2020

```{r include=FALSE}
# install.packages("librarian")
librarian::shelf(tidyverse, ggplot2, reshape2, ggtext, jsonlite, stringr, janitor, tools, fs)
```

```{r function-definitions, echo=FALSE, message=FALSE}
wrangle <- function(filename_root){
  jsonl_file <- paste(filename_root, ".jsonl", sep="")
  lines <- readLines(jsonl_file)
  
  config <- lapply(lines[1], fromJSON)  %>% 
  #  # flatten the items
    lapply(unlist) %>% 
    bind_rows() %>%
    clean_names()
  #config <- fromJSON(lines[1], flatten=TRUE) %>% 
    # extract the first element of the list (what an ugly expression)
  #  `[[`(1)
  
  csv <- read_csv(file=str_glue("{filename_root}.csv", sep="",  show_col_types = FALSE)) %>%
    select(-last_col()) %>%
    clean_names() %>%
    as_tibble() %>%
    tibble::rowid_to_column("index") 
  
  # convert to a long form
  tb <- pivot_longer(csv %>% 
                       select(c(index, 
                                mean_vehicle_count, 
                                vehicle_fraction_idle,
                                vehicle_fraction_picking_up, 
                                vehicle_fraction_with_rider,
                                mean_trip_wait_fraction,
                                mean_trip_wait_time)), 
                     cols = c("vehicle_fraction_idle", 
                              "vehicle_fraction_picking_up", 
                              "vehicle_fraction_with_rider", 
                              "mean_trip_wait_fraction",
                              "mean_trip_wait_time"),
             names_to="measure",
             values_to="value"
             ) 
  return (list(tb=tb, config=config))
}

sequence_plot <- function(tb, config){
  annotation = str_glue("City size={config$config_city_size} blocks<br>
                        Request rate={config$config_base_demand} per block<br>
                        Mean trip distance=20 minutes<br>
                        Trip inhomogeneity={config$config_trip_inhomogeneity}")
  colours <- c("vehicle_fraction_idle" = "royalblue3", 
               "vehicle_fraction_picking_up" = "gold3", 
               "vehicle_fraction_with_rider" = "green4", 
               "mean_trip_wait_fraction" = "red3")
  linetypes <-  c("vehicle_fraction_idle" = "solid", 
                  "vehicle_fraction_picking_up" = "solid", 
                  "vehicle_fraction_with_rider" = "solid", 
                  "mean_trip_wait_fraction" = "dashed")
  labels <- c("vehicle_fraction_idle" = "P1 (available)",
              "vehicle_fraction_picking_up" = "P2 (dispatch)",
              "vehicle_fraction_with_rider" = "P3 (busy)",
              "mean_trip_wait_fraction" = "Wait fraction (W/L)")
  
  fit_cutoff <- tb %>% 
    filter(measure=="vehicle_fraction_idle") %>%
    filter(value == 0) %>%
    summarise(max(mean_vehicle_count)) %>%
    pull()

  p <- ggplot(data=tb, 
              mapping=aes(x=mean_vehicle_count, 
                          y=value, 
                          colour=measure))
  p + 
    geom_point(alpha=0.6, size=2) +
    geom_line(stat="smooth", 
              data=subset(tb, mean_vehicle_count >= fit_cutoff), 
              alpha=0.6, size=1) +
    scale_x_continuous(limits=c(round(min(tb$mean_vehicle_count)),
                                round(max(tb$mean_vehicle_count)))) +
    scale_y_continuous(limits=c(0,1), minor_breaks = seq(0 , 1, .05), breaks = seq(0, 1, .2)) +
    scale_linetype_manual(values=linetypes, labels=labels) +
    scale_colour_manual(values=colours, labels=labels) + 
    labs(title=config$config_title,
         x="Vehicles",
         y="Fraction") +
    geom_richtext(aes(x=round(max(mean_vehicle_count)) * 0.7, 
                      y=1.0, label=annotation), #width=unit(0.3, "npc"), stat="unique"),
                 size=3.2, 
                 lineheight = .9,
                 color="black", 
                 fill="white", 
                 hjust = "left",
                 vjust= "top",
                 label.colour = "white",
                 show.legend = FALSE) +
    theme(legend.position = c(0.2, 0.85), 
          legend.background = element_rect(fill=alpha("lightgray", 0),
                                    size=0.5, linetype="blank", 
                                    colour ="lightgray"),
          legend.key=element_rect(fill = alpha("white", 0)),
          legend.title = element_blank(), 
          legend.margin=margin(t=-0.25,l=0.05,b=0.0,r=0.05, unit='cm'), 
          legend.spacing.y = unit(0.5, "lines")
          )
  # return (longseq)
}
```


```{r plot-sequence, echo=FALSE}
filename_root="./toronto-feb-6-2020-sequence"
# filename_root="./scarborough-seq-2022-07-24-11-35"
# filename_root="./scarborough-seq-3-2022-07-24-11-41"
#filename_root="./wait_times/homog_32-r5-2022-07-30-13-44"
x <- wrangle(filename_root = filename_root)
tb <- x$tb
config <- x$config
sequence_plot(tb=tb, config=config)
# sequence_plot(tb=tb, config=config, fit_cutoff = 10)
```

```{r wait-time-focus, echo=FALSE}
p <- ggplot(data=tb %>% filter(measure=="mean_trip_wait_time"), 
            mapping=aes(x=mean_vehicle_count, y=value))

p +
  geom_point(colour="darkred", alpha=0.6, size=2) +
  geom_line(stat="smooth", linetype="dashed", colour="darkred", alpha=0.6, size=.5) +
  scale_x_continuous() +
  scale_y_continuous(limits=c(0, 10)) +
  labs(title="Vehicles needed to provide a given wait time (135 trips/min)",
       x="Number of vehicles",
       y="Wait time (minutes)")
```

# Uber's business depends on high demand

## Maximum utilization rates

Create a data frane of P3_max and N_min values, using a 32*32 city. Cities with high demand can support a higher P3_max. I am surprised that the number of vehicles per trip required is c    onstant (that is, N_min = a R + b).

What is *M_w=0.30*? It must be the vale

```{r p3-vs-demand, echo=FALSE}
labels <- c("0.0" = "0.0", "0.5" = "0.5")

tb <- read_csv("demand.csv", col_types="nnccn") %>% 
  as_tibble()

p<- ggplot(data=tb %>% filter(variable %in% c("P3_max", "N_min")), 
           mapping=aes(x=demand, y=value, linetype=inhomogeneity, colour=variable))
p + 
  geom_line(alpha=0.7, size=1) +
  geom_point(alpha=0.7, size=3) +
  ggtitle("Maximum utilization rates",
          subtitle="Minimum vehicle count and max utilization for a 32*32 city") +
  scale_colour_manual(labels=labels) + 
  # theme(legend.position = "none") +
  scale_color_brewer(palette= "Dark2") + 
  facet_wrap(~variable, scales = "free") +
  xlim(0,50) 
```


Fitting with a polynomial

```{r polyfit, echo=FALSE}
model <- lm(tb %>% filter(variable=='N_min', inhomogeneity=='0.0'), formula=value ~ demand + I(demand^2))
summary(model)
```

## How much commission can Uber take?

At different levels of demand, the commission Uber can take to support a given wait time. This uses a 32*32 city, with a wait time of 0.2 (this would be W/L = 0.2, so if a block is a minute, and <L> is 16 minutes, then W=3.2 minutes).

The sequences are run with different demand, but everything else the same, including equilibration. As a result, the P3 values are the same across all demand levels (because the income is the reservation wage).

Settings are: price=1.5, commission (m) varies, demand elasticity = 0, reservation wage = 0.5.


```{r commission-3min, echo=FALSE}
p<- ggplot(data=tb %>% filter(variable %in% c("M_w=0.30", "M_w=0.25")), 
           mapping=aes(x=demand, y=value, linetype=inhomogeneity, colour=variable))

p + 
  geom_point(alpha=0.8, size=3) +
  geom_smooth(method = "glm", formula = y ~ log(x), alpha=0.8, se=FALSE) +
  ggtitle("Commission yielding a wait fraction of 0.3 (4.8mins)",
          subtitle="At high demand, and in homogeneous environments,\nthe platform can take a higher commission and still deliver the same wait time") +
  xlab("Demand (requests/block)") +
  ylab("Platform commission") +
  scale_color_brewer(palette= "Dark2") 
```

## How much does P3 increase with demand?

P3 for a given wait time W increases with demand, showing the possibility of decreasing W while increasing P3, which is 
one of Uber's central claims. But there is no quantitative statement of this - it's an elasticity of a sort. This chart
shows P3 vs demand D for a Toronto Feb 6 2020 city (C=60, I=0.4, L=[0,40]).

```{r p3_versus_d}
labels <- c("N_w=0.30" = "0.0", 
            "0.5" = "N_w=0.30")

tb <- read_csv("demand.csv", col_types="nnccn") %>% 
  as_tibble() %>% 
  filter(variable %in% c("P3_w=0.30", "N_w=0.30"))

p<- ggplot(data=tb , 
           mapping=aes(x=demand, y=value, colour=variable))
p + 
  geom_point(alpha=0.7, size=2) +
  geom_smooth(method="glm", formula = y ~ poly(x, 2), se = FALSE, size=1, alpha=0.7) +
  facet_wrap(~variable, scales = "free") +
  scale_x_continuous(limits=c(0, NA)) +
  scale_y_continuous(limits=c(0, NA)) +
  ggtitle("Utilization rates depend on demand",
          subtitle="P3 for W=0.3 in Toronto") +
  scale_colour_manual(labels=labels) + 
  theme(legend.position = "none") +
  scale_color_brewer(palette= "Dark2")
```

# Wait times

```{r prep-wait-times, echo=FALSE, message=FALSE}
#csv_files <- Sys.glob(file.path("./wait_times", "*.csv"))
csv_files <- Sys.glob(file.path("./wait_times", "*.csv"))

configs = NULL
sequences = NULL
for(csv_file in csv_files){
  # print(sprintf("csv_file=%s", csv_file))
  file_path_sans_ext <- tools::file_path_sans_ext(csv_file)
  sequence_id <- path_file(file_path_sans_ext)
  jsonl_file <- sprintf("%s.jsonl", file_path_sans_ext)
  lines <- readLines(jsonl_file)
  # print(sprintf("jsonl_file=%s", jsonl_file))
  
  # assemble the config elements into a tibble
  if(is.null(configs)){
    configs <- lapply(lines[1], fromJSON)  %>% 
      #  # flatten the items
      lapply(unlist) %>% 
      bind_rows() %>%
      clean_names() %>%
      as_tibble() %>%
      mutate(sequence_id=sequence_id,
             config_base_demand=as.integer(config_base_demand))
    
    sequences <- read_csv(file=csv_file, show_col_types = FALSE) %>%
      select(-last_col()) %>%
      clean_names() %>%
      as_tibble() %>%
      mutate(sequence_id=sequence_id) %>%
      tibble::rowid_to_column("index") %>%
      select(c(sequence_id,
               index, 
               mean_vehicle_count, 
               vehicle_fraction_idle,
               vehicle_fraction_picking_up, 
               vehicle_fraction_with_rider,
               mean_trip_wait_fraction,
               mean_trip_wait_time)) %>%
      pivot_longer(cols = c("vehicle_fraction_idle", 
                          "vehicle_fraction_picking_up", 
                          "vehicle_fraction_with_rider", 
                          "mean_trip_wait_fraction",
                          "mean_trip_wait_time"),
                       names_to="measure",
                       values_to="value")
  } else {
    this_config <- lapply(lines[1], fromJSON)  %>% 
      #  # flatten the items
      lapply(unlist) %>% 
      bind_rows() %>%
      clean_names() %>%
      mutate(sequence_id=sequence_id,
             config_base_demand=as.integer(config_base_demand))
    
    configs <- union(configs, this_config)
    
    this_sequence <- read_csv(file=csv_file, show_col_types = FALSE) %>%
      select(-last_col()) %>%
      clean_names() %>%
      as_tibble() %>%
      mutate(sequence_id=sequence_id) %>%
      tibble::rowid_to_column("index") %>%
      select(c(sequence_id,
               index, 
               mean_vehicle_count, 
               vehicle_fraction_idle,
               vehicle_fraction_picking_up, 
               vehicle_fraction_with_rider,
               mean_trip_wait_fraction,
               mean_trip_wait_time)) %>%
      pivot_longer(cols = c("vehicle_fraction_idle", 
                            "vehicle_fraction_picking_up", 
                            "vehicle_fraction_with_rider", 
                            "mean_trip_wait_fraction",
                            "mean_trip_wait_time"),
                   names_to="measure",
                   values_to="value")
    sequences <- union(sequences, this_sequence)
  }
}
```

```{r plot-wait-times, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
tb <- sequences %>% 
  filter(measure=="mean_trip_wait_time") %>%
  left_join(configs, by="sequence_id") %>%
  select(c(config_base_demand, config_trip_inhomogeneity, mean_vehicle_count, measure, value)) %>%
  mutate(vehicles_per_trip = mean_vehicle_count / config_base_demand) %>%
  filter(config_trip_inhomogeneity > 0.0)

p <- ggplot(tb,
            mapping=aes(x=value,
                        y=vehicles_per_trip,
                        colour=factor(config_base_demand)))

p +
  geom_point(size=1, alpha=0.6) +
  geom_smooth(method = "lm", formula=y ~ I(1/x), alpha=0.8, se=F) +
  #facet_wrap(~config_trip_inhomogeneity) +
  # geom_line( size=1.5, alpha=0.6) +
  scale_y_continuous(limits=c(0,200)) +
  scale_x_continuous(limits=c(1.5,6)) +
  labs(title="Wait times and vehicle counts",
       subtitle="City size=32, trip inhomogeneity = 0.4",
       x="Wait time (minutes)",
       y="Vehicles per unit demand (trips per minute)") +
  guides(size=FALSE) +
  guides(colour=guide_legend(title="Demand\n(Trips per minute)")) 
```
