---
title: 'Ridehail Notebooks: Sequences'
output:
  html_document:
    df_print: paged
---

# Toronto, Feb 6, 2020

```{r include=FALSE}
# install.packages("librarian")
librarian::shelf(tidyverse, ggplot2, reshape2, ggtext, jsonlite, stringr)
```

```{r echo=FALSE}
sequence_plot <- function(filename_root, fit_cutoff){
  jsonl_file <- paste(filename_root, ".jsonl", sep="")
lines <- readLines(jsonl_file)
config <- lapply(lines[1], fromJSON)
config <- lapply(config, unlist)
config <- bind_rows(config)
annotation = str_glue("City size={config$config.city_size} blocks<br>
                      Request rate={config$config.base_demand} per block<br>
                      Mean trip distance=20 minutes<br>
                      Trip inhomogeneity={config$config.trip_inhomogeneity}")
sequence <- read_csv(file=str_glue("{filename_root}.csv", sep=""))
sequence$idx <- 1:nrow(sequence)
# convert to a long form
longseq <- melt(sequence, 
                measure.vars = c("vehicle_fraction_idle", 
                                 "vehicle_fraction_picking_up", 
                                 "vehicle_fraction_with_rider", 
                                 "mean_trip_wait_fraction")
                )
  colours <- c("vehicle_fraction_idle" = "royalblue3", "vehicle_fraction_picking_up" = "gold3", "vehicle_fraction_with_rider" = "green4", "mean_trip_wait_fraction" = "red3")
  linetypes <-  c("vehicle_fraction_idle" = "solid", "vehicle_fraction_picking_up" = "solid", "vehicle_fraction_with_rider" = "solid", "mean_trip_wait_fraction" = "dashed")
  labels <- c("vehicle_fraction_idle" = "P1 (available)", "vehicle_fraction_picking_up" = "P2 (dispatch)", "vehicle_fraction_with_rider" = "P3 (busy)", "mean_trip_wait_fraction" = "Wait fraction (W/L)")
  p <- ggplot(data=longseq, mapping=aes(x=mean_vehicle_count, y=value, colour=variable, linetype=variable, label=annotation))
  p + geom_point(alpha=0.5, size=2) +
    geom_line(stat="smooth", 
              data=subset(longseq, mean_vehicle_count > fit_cutoff), 
              alpha=0.5, size=1) +
    scale_x_continuous(minor_breaks = seq(0, 10000, 500), breaks = seq(0, 10000, 1000)) +
    scale_y_continuous(limits=c(0,1), minor_breaks = seq(0 , 1, .05), breaks = seq(0, 1, .2)) +
    scale_linetype_manual(values=linetypes, labels=labels) +
    scale_colour_manual(values=colours, labels=labels) + 
    xlab("Vehicles") +
    ylab("Fraction") + 
    ggtitle(config$config.title) +
    geom_richtext(aes(x=8000, y=1.0, label=annotation, width=unit(0.3, "npc"), stat="unique"),
                 size=3.2, 
                 lineheight = .9,
                 color="black", 
                 fill="white", 
                 hjust = "left",
                 vjust= "top",
                 label.colour = "white",
                 show.legend = FALSE) +
    # theme(axis.line = element_line(colour="darkgrey", size=1))
    theme(legend.position = c(0.2, 0.85), 
          legend.background = element_rect(fill=alpha("lightgray", 0),
                                    size=0.5, linetype="blank", 
                                    colour ="lightgray"),
          legend.key=element_rect(fill = alpha("white", 0)),
          legend.title = element_blank(), 
          legend.margin=margin(t=-0.25,l=0.05,b=0.0,r=0.05, unit='cm'), 
        legend.spacing.y = unit(0.5, "lines")) 
}
```


```{r echo=FALSE}
filename_root="./toronto-feb-6-2020-sequence"
sequence_plot(filename_root = filename_root, fit_cutoff = 4500)
```

# Uber's business depends on high demand

## Maximum utilization rates

Create a data frane of P3_max and N_min values, using a 32*32 city. Cities with high demand can support a higher P3_max. I am surprised that the number of vehicles per trip required is constant (that is, N_min = a R + b).

```{r echo=FALSE}
df <- data.frame( city_size=numeric(0), demand=numeric(0), inhomogeneity=character(0), variable=character(0), value=numeric(0))
df <- read_csv("demand.csv", col_types="nnccn")
tb <- as_tibble(df)
colours <- c("P3_max"="royalblue3", "N_min"="deeppink3", "M_w=0.30"="sienna3" )
labels <- c("0.0" = "0.0", "0.5" = "0.5")
p<- ggplot(data=tb %>% filter(variable %in% c("P3_max", "N_min")), 
           mapping=aes(x=demand, y=value, linetype=inhomogeneity, colour=variable))
p + geom_line(alpha=0.7, size=1) +
  geom_point(alpha=0.7, size=3) +
  scale_colour_manual(values=colours, labels=labels) + 
  # theme(legend.position = "none") +
  # scale_color_brewer(palette= "Dark2") + 
  facet_wrap(~variable, scales = "free")
```

## How much commission can Uber take?

At different levels of demand, the commission Uber can take to support a given wait time. This uses a 32*32 city, with a wait time of 0.2 (this would be W/L = 0.2, so if a block is a minute, and <L> is 16 minutes, then W=3.2 minutes).

The sequences are run with different demand, but everything else the same, including equilibration. As a result, the P3 values are the same across all demand levels (because the income is the reservation wage).

Settings are: price=1.5, commission (m) varies, demand elasticity = 0, reservation wage = 0.5.


```{r}
p<- ggplot(data=tb %>% filter(variable %in% c("M_w=0.30")), 
           mapping=aes(x=demand, y=value, linetype=inhomogeneity, colour=variable))
p + 
  # geom_line(alpha=0.7, size=1) +
  geom_point(alpha=0.8, size=3) +
  geom_smooth(method = "gam", formula = y ~ log(x), alpha=0.8) +
  ggtitle("Commission yielding a wait fraction of 0.3 (4.8mins)",
          subtitle="At high demand, and in homogeneous environments,\nthe platform can take a higher commission and still deliver the same wait time") +
  xlab("Demand (requests/block)") +
  ylab("Platform Commission") +
  scale_colour_manual(values=colours, labels=labels)
  # theme(legend.position = "none") +
  # scale_color_brewer(palette= "Dark2") + 
  #facet_wrap(~variable, scales = "free")
```